Sound and Automated Synthesis of Digital Controllers for Physical Plants

### 2016-10-06 paper/paper.tex
hscc17

\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\lstset{
  mathescape, 
  language={C},
  basicstyle=\small
}
\usepackage{enumitem}      % adjust spacing in enums
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   
\usepackage{graphicx}
\usepackage{float,subfig}
\usepackage{xspace,framed}
\usepackage{colortbl}
\usepackage{calc}
\usepackage[dvipsnames]{xcolor}

@@

\newtheorem{myassumption}{Assumption}
\newtheorem{mylemma}{Lemma}

@@

Consider the closed-loop sampled system in Figure~\ref{fig:sampledsystem},
where $C(z)$ is the digital controller model, whose control output signal is
quantized ($Q2$) and interpolated via a ZOH using digital-to-analog
converter (DAC).  $G(s)$ is the continuous-time model of the plant, from
which output samples are taken and quantized ($Q1$) by means of an
analog-to-digital converter (ADC) synchronized with the DAC.  Each quantizer
($Q1$ and $Q2$) will present some contribution for the total quantization
noise.
%
\begin{figure}[ht]
\centering
\resizebox{.5\textwidth}{!}{
 \begin{tikzpicture}[scale=0.3,-,>=stealth',shorten >=.2pt,auto,
     semithick, initial text=, ampersand replacement=\&,]

@@

First, some assumptions have to be established to obtain a model for $\Delta{G_{Q}(z)}$. 
\blue{[Here we need to be careful: we first argue that quantisation (spelled with an s, so that ER II is happy) effects can be represented as an additive noise (great), which however is a stochastic signal. But we handle the noise deterministically. How so? Where is the catch? ]}
%
\begin{myassumption}
\label{whitenoise}
%
The quantization noises $\nu_{1}$ (from the $Q1$) and $\nu_{2}$ (from the
$Q2$) \red{[do we need the second? See my further comments below.]} \green{There was a typo in \eqref{eq:deltaG_final}. It was missing $q_{1}$ and $q_{1}$, both effects are considered here. }
are white noises.
%
\end{myassumption}

@@

\begin{myassumption}
%
The controller and plant are linear, {\it i.e.} any source of nonlinearity ({\it e.g.} overflows) may be disregard. %\red{[explain difference btw overflows and saturations. Also, do we care about limit cycles? I'd quite happily disregard them.]}
%
\end{myassumption}

@@

Given those assumptions, the present approach is further based on the following lemmas. 
\red{[So we have introduced two quantisation errors, properly indexed. Next we discuss a non-indexed entity - pls clarify or comment.]}
%
\begin{mylemma}
%
\cite{widrow1956} The second moment (variance) of the probability probability density function of quantization
noise is given by %\textcolor{red}{expand the acronym PDF}
%
\begin{equation}
\label{eq:variancelemma}
\sigma_{Q}^{2}=\frac{q^{2}}{12},
\end{equation}
\end{mylemma}

\noindent where $q$ is the step size of quantizer.

\begin{mylemma} \cite{astrom1997computer}
%
The pulse transfer function $G(z)$ is a z-domain model that relates the output sampled by a ADC and the digital input provided by ZOH. It is equivalent at the sample times ({\it i.e.} $kT$ to the continuous transfer function $G(s)$ driven by ZOH.  This pulse transfer
function can be computed as
%
\begin{equation}
\label{eq:pulsetf}
G(z)=(1-z^{-1})\mathcal{Z}\left\lbrace{\mathcal{L}^{-1}\left\lbrace{\frac{G(s)}{s}}\right\rbrace_{t=kT}}\right\rbrace.
\end{equation}
\end{mylemma}
\red{[Here the dependence on the sampling time T needs to appear somewhere.]} \green{It appears,  $kT$ is the time $t$ used after the inverse Laplace transform}

Considering the quantization noise $\nu(k)$ as an output error in the
discrete plant model, {\it i.e.}, it does not affect the plant dynamics, but
its output samples, it is easy to prove that $\Delta{G_{Q}(z)}$ can be
obtained from
%
\begin{equation}
\label{eq:quantization_tf}
\Delta{G_{Q}(z)}=\frac{\nu(z)}{U(z)},
\end{equation}

\noindent where $\nu(z)$ is the z-transform of the random process $\nu(k)$,
which can be calculated through the power spectral density of
$\nu(k)$~\cite{poularikas2000transforms}.  Thus, the $\Delta{G_{Q}(z)}$ can
expressed as
%
\begin{equation}
\label{eq:deltag_var}
\Delta{G_{Q}(z)}=\frac{\sigma_{Q}^{2}}{U(z)},
\end{equation}

\noindent where $\sigma_{Q}^{2}$ is the variance of the quantization noise and
$U(z)$ is the z-transform of control signal ({\it i.e.}, the z-transform of
the signal from the digital controller), which is provided to the plant by
means of ZOH.

\red{[Please write out a formal statement for the proof below:]}
%
\begin{proof}
%
Suppose that the plant discrete model in Figure~\ref{fig:sampledsystem} can
be represented by the following pulse transfer function (without
quantization noise effect):
%
\begin{equation}
\label{eq:tfwithout}
G(z)=\frac{Y(z)}{U(z)}=\frac{B(z)}{A(z)}=\frac{b_{0}+b_{1}z^{-1}+b_{2}z^{-2}+\cdots+b_{M}z^{-M}}{1+a_{1}z^{-1}+a_{2}z^{-2}+\cdots+a_{N}z^{-N}},
\end{equation}
%
\noindent where $Y(z)$ is the z-transform of the plant output signal without
noise $y(k)$, $U(z)$ is the z-transform of the control signal, which is
provided to the plant, $A(z)$ and $B(z)$ are the plant pulse transfer
function numerator and denominator, respectively.

>>>>

\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\lstset{
  mathescape, 
  language={C},
  basicstyle=\small
}
\usepackage{enumitem}      % adjust spacing in enums
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   
\usepackage{graphicx}
\usepackage{float,subfig}
\usepackage{xspace,framed}
\usepackage{colortbl}
\usepackage{calc}
\usepackage[dvipsnames]{xcolor}
\usepackage{amssymb,amsmath,amsfonts} 
\usepackage{mathtools}

@@

\newtheorem{myassumption}{Assumption}
\newtheorem{mylemma}{Lemma}
\newtheorem{myprop}{Proposition}

@@

Consider the closed-loop sampled system in Figure~\ref{fig:sampledsystem} \red{[requires updating]},
where $C(z)$ is the digital controller model, whose control output signal is
quantized ($Q2$) and interpolated via a ZOH using digital-to-analog
converter (DAC).  
\red{[do we really need to regard the quantisation effects of the ZOH in the DAC? Elaborate next:]}
$G(s)$ is the continuous-time model of the plant, from
which output samples are taken and quantized ($Q1$) by means of an
analog-to-digital converter (ADC) synchronized with the DAC.  Each quantizer
($Q1$ and $Q2$) will present some contribution for the total quantization
noise.
%
\begin{figure}[ht]
\centering
\resizebox{.5\textwidth}{!}{
 \begin{tikzpicture}[scale=0.3,-,>=stealth',shorten >=.2pt,auto,
     semithick, initial text=, ampersand replacement=\&,]

@@

First, some assumptions have to be established to obtain a model for $\Delta{G_{Q}(z)}$. 
\blue{[Here we need to be careful: we first argue that quantisation (spelled with an s, so that ER II is happy) effects can be represented as an additive noise (great), which however is a stochastic signal. But we handle the noise deterministically. How so? Where is the catch? ]}
%
\begin{myassumption}
\label{whitenoise}
%
The quantization noises $\nu_{1}$ (from the $Q1$) and $\nu_{2}$ (from the
$Q2$) \red{[do we need the second? See my further comments below.]} \green{There was a typo in \eqref{eq:deltaG_final}. It was missing $q_{1}$ and $q_{1}$, both effects are considered here. } \red{[why are both needed? I'm still unconvinced.]}
are white noises.
%
\end{myassumption}

@@

\begin{myassumption}
%
The controller and the plant are linear, {\it i.e.} any source of nonlinearity ({\it e.g.} overflows) may be disregard. %\red{[explain difference btw overflows and saturations. Also, do we care about limit cycles? I'd quite happily disregard them.]}
%
\end{myassumption}

@@

Given those assumptions, the present approach is further based on the following lemmas. 
\red{[So we have introduced two quantisation errors, properly indexed. Next we discuss a non-indexed entity - pls clarify or comment.]}
%
\begin{mylemma}
%
\cite{widrow1956} The second moment (variance) of the probability probability density function of quantization
noise is given by %\textcolor{red}{expand the acronym PDF}
%
\begin{equation}
\label{eq:variancelemma}
\sigma_{Q}^{2}=\frac{q^{2}}{12},
\end{equation}
\end{mylemma}
where $q$ is the step size of quantizer. \blue{[can you add a comment on this assumption? Why the factor 12?]}

\begin{mylemma} \cite{astrom1997computer}
%
The pulse transfer function $G(z)$ is a z-domain model that relates the output sampled by a ADC and the digital input provided by ZOH. It is equivalent at the sample times ({\it i.e.} $kT$ to the continuous transfer function $G(s)$ driven by ZOH.  This pulse transfer
function can be computed as
%
\begin{equation}
\label{eq:pulsetf}
G(z) = (1-z^{-1})\mathcal{Z}\left\lbrace{\mathcal{L}^{-1}\left\lbrace{\frac{G(s)}{s}}\right\rbrace_{t=kT}}\right\rbrace.
\end{equation}
\end{mylemma}
\red{[Here the dependence on the sampling time T needs to appear somewhere.]} 
\green{It appears,  $kT$ is the time $t$ used after the inverse Laplace transform}
\red{[ok, then let's call it $G(z,T)$ as I suggested earlier.]}

\begin{myprop}
Considering the quantisation noise $\nu(k)$ as an output error in the
discrete plant model, 
%{\it i.e.}, it does not affect the plant dynamics, but its output samples, 
$\Delta{G_{Q}(z)}$ can be obtained from
%
\begin{equation}
\label{eq:quantization_tf}
\Delta{G_{Q}(z)}=\frac{\nu(z)}{U(z)},
\end{equation}
where $\nu(z)$ is the z-transform of the random process $\nu(k)$,
which can be calculated through the power spectral density of
$\nu(k)$~\cite{poularikas2000transforms}.  
Thus, the $\Delta{G_{Q}(z)}$ can expressed as
%
\begin{equation}
\label{eq:deltag_var}
\Delta{G_{Q}(z)}=\frac{\sigma_{Q}^{2}}{U(z)},
\end{equation}
where $\sigma_{Q}^{2}$ is the variance of the quantization noise and
$U(z)$ is the z-transform of control signal ({\it i.e.}, the z-transform of
the signal from the digital controller), which is provided to the plant by
means of ZOH.
\end{myprop}

%
\begin{proof}
%
Suppose that the plant discrete model in Figure~\ref{fig:sampledsystem} can
be represented by the following pulse transfer function (without
quantization noise effect):
%
\begin{equation}
\label{eq:tfwithout}
G(z)=\frac{Y(z)}{U(z)}=\frac{B(z)}{A(z)}=\frac{b_{0}+b_{1}z^{-1}+b_{2}z^{-2}+\cdots+b_{M}z^{-M}}{1+a_{1}z^{-1}+a_{2}z^{-2}+\cdots+a_{N}z^{-N}},
\end{equation}
%
\noindent where $Y(z)$ is the z-transform of the plant output signal without
noise $y(k)$, $U(z)$ is the z-transform of the control signal, which is
provided to the plant, $A(z)$ and $B(z)$ are the plant pulse transfer
function numerator and denominator, respectively.
